---
title: "hw2-02"
author: "shreyakolte"
format: html
editor: visual
---

# IMPUTING LIKE A DATA SCIENTIST:

### Required Setup:

First we will install all the packages required for our operations on our dataset.

```{r}
# Sets the number of significant figures to two - e.g., 0.01
options(digits = 2)

# Required package for quick package downloading and loading 
if (!require(pacman))
  install.packages("pacman")

pacman::p_load(cluster, # K cluster analyses
               dlookr, # Exploratory data analysis
               formattable, # HTML tables from R outputs
               ggfortify, # Plotting tools for stats
               ggpubr, # Publishable ggplots
               here, # Standardizes paths to data
               kableExtra, # Alternative to formattable
               knitr, # Needed to write HTML reports
               missRanger, # To generate NAs
               plotly, # Visualization package
               rattle, # Decision tree visualization
               rpart, # rpart algorithm
               tidyverse, # Powerful data wrangling package suite
               visdat) # Another EDA visualization package

# Set global ggplot() theme
# Theme pub_clean() from the ggpubr package with base text size = 16
theme_set(theme_pubclean(base_size = 16)) 
# All axes titles to their respective far right sides
theme_update(axis.title = element_text(hjust = 1))
# Remove axes ticks
theme_update(axis.ticks = element_blank()) 
# Remove legend key
theme_update(legend.key = element_blank())


```

### LOADING AND DOWNLOADING THE DATASET:

Here we will load the dataset which we are going to use for doing all our operations.

```{r}
# Let's load a data set from the diabetes data set
demographics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-09-05/demographics.csv')
demographics


# What does the data look like?
demographics |>
  head() |>
  formattable()
```

### DIAGNOSE YOUR DATA:

The data set which we are using is being diagnosed here.

```{r}
# What are the properties of the data
demographics |>
  diagnose() |>
  formattable()
```

### DIAGNOSE OUTLIERS:

Here we will diagnose the outliers by plotting a box plot and histogram.

```{r}
# Table showing outliers
demographics |>
  diagnose_outlier() |>
  filter(outliers_ratio > 0) |>  
  mutate(rate = outliers_mean / with_mean) |>
  arrange(desc(rate)) |> 
  select(-outliers_cnt) |>
  formattable()
```

```{r}
# Boxplots and histograms of data with and without outliers
demographics |>
  select(find_outliers(demographics)) |>
           plot_outlier()
```

## **Basic Exploration of Missing Values (NAs):**

Table showing the extent of NAs in columns containing them

```{r}
# Randomly generate NAs for 30
na.demographics <- demographics |>
  generateNA(p = 0.3)

# First six rows
na.demographics |>
  head() |>
  formattable()
```

```{r}
# Create the NA table
na.demographics |>
  plot_na_pareto(only_na = TRUE, plot = FALSE) |>
  formattable() # Publishable table
```

-   Plots showing the frequency of missing values

    ```{r}
    # Plot the insersect of the columns with missing values
    # This plot visualizes the table above
    na.demographics |>
      plot_na_pareto(only_na = TRUE)
    ```

## **Advanced Exploration of Missing Values (NAs):**

Intersect plot that shows, for every combination of columns relevant, how many missing values are common

Orange boxes are the columns in question

x axis (top green bar plots) show the number of missing values in that column

y axis (right green bars) show the number of missing values in the columns in orange blocks

```{r}
# Plot the intersect of the 5 columns with the most missing values
# This means that some combinations of columns have missing values in the same row
na.demographics |>
  select(covered,employment,year) |>
  plot_na_intersect(only_na = TRUE) 
```

### **Determining if NA Observations are the Same:**

Missing values can be the same observation across several columns. The visdat package can solve this with the `vis_miss()` function which shows the rows with missing values through `ggplotly()`

```{r}
# Interactive plotly() plot of all NA values to examine every row
#na.demographics |>
 #select(covered, employment, year) |>
 #vis_miss() |>
 #ggplotly() 
```

### **Classifying Outliers:**

Before imputing outliers, you will want to diagnose whether it's they are natural outliers or not. 

```{r warning=FALSE}
# Box plot
demographics %>% # Set the simulated normal data as a data frame
  ggplot(aes(x =year, y = facet, fill =facet)) + # Create a ggplot
  geom_boxplot(width = 0.5, outlier.size = 2, outlier.alpha = 0.5) +
  xlab("Year") +  # Relabel the x axis label
  ylab("Covered") + # Remove the y axis label
  theme(legend.position = "none")  # Remove the legend 
```

### Mean Imputation:

The mean of the observed values for each variable is computed and the outliers for that variable are imputed by this mean

```{r}
# Raw summary, output suppressed
mean_out_imp_p_covered <- demographics |>
  select(p_covered) |>
  imputate_outlier(p_covered, method = "mean")

# Output showing the summary statistics of our imputation
mean_out_imp_p_covered |>
  summary() 
```

```{r}
# Visualization of the mean imputation
mean_out_imp_p_covered |>
  plot()
```

### Median Imputation:

The median of the observed values for each variable is computed and the outliers for that variable are imputed by this median

```{r}
# Raw summary, output suppressed
med_out_imp_p_covered <- demographics |>
  select(p_covered) |>
  imputate_outlier(p_covered, method = "median")

# Output showing the summary statistics of our imputation
med_out_imp_p_covered |>
  summary()
```

```{r}
# Visualization of the median imputation
med_out_imp_p_covered |>
  plot()
```

### Mode Imputation:

The mode of the observed values for each variable is computed and the outliers for that variable are imputed by this mode

```{r}
# Raw summary, output suppressed
mode_out_imp_p_covered <- demographics |>
  select(p_covered) |>
  imputate_outlier(p_covered, method = "mode")

# Output showing the summary statistics of our imputation
mode_out_imp_p_covered |>
  summary()
```

```{r}
# Visualization of the mode imputation
mode_out_imp_p_covered |>
plot()
```

### **Capping Imputation (aka Winsorizing):**

The Percentile Capping is a method of Imputing the outlier values by replacing those observations outside the lower limit with the value of 5th percentile and those that lie above the upper limit, with the value of 95th percentile of the same dataset.

```{r}
# Raw summary, output suppressed
cap_out_imp_p_covered <- demographics |>
  select(p_covered) |>
  imputate_outlier(p_covered, method = "capping")

# Output showing the summary statistics of our imputation
cap_out_imp_p_covered |>
  summary()
```

```{r}
# Visualization of the capping imputation
cap_out_imp_p_covered |>
  plot()
```

## **Imputing NAs:**

We can use the impute_na() method.

### **K-Nearest Neighbor (KNN) Imputation:**

KNN is a machine learning algorithm that classifies data by similarity. This in effect clusters data into similar groups. The algorithm predicts values of new data to replace NA values based on how closely they resembles training data points, such as by comparing across other columns.

```{r}
# KNN plot of our dataset without categories
data_omit <- na.omit(demographics)
demographicsAlt <- demographics
demographicsAlt <- top_n(demographics,20,facet)
autoplot(clara(data_omit[-5], 3))



```

```{r}
#demographics.noNA
```

```{r}
# Raw summary, output suppressed
#demographicsnona <- demographics.noNA 
#knn_na_imp_pa_covered <- demographicsnona|>
# imputate_na(covered, method = "knn")

# Plot showing the results of our imputation
#knn_na_imp_pa_covered |>
# plot()
```

#### There are no missing values in the dataset.

### **Recursive Partitioning and Regression Trees (rpart):**

rpart is a decision tree machine learning algorithm that builds classification or regression models through a two stage process, which can be thought of as binary trees. The algorithm splits the data into subsets, which move down other branches of the tree until a termination criteria is reached.

```{r}
library(magrittr)
# Raw summary, output suppressed
non_numeric <- demographics %>%
  select_if(is.numeric)

rpart_na_imp_p_covered <- non_numeric %>%
  imputate_na(p_covered, method = "rpart")

# Plot showing the results of our imputation
rpart_na_imp_p_covered %>%
  plot()
```

### **Multivariate Imputation by Chained Equations (MICE):**

MICE is an algorithm that fills missing values multiple times, hence dealing with uncertainty better than other methods. This approach creates multiple copies of the data that can then be analyzed and then pooled into a single dataset

```{r warning=FALSE}
# Raw summary, output suppressed
mice_na_imp_p_covered <- na.demographics |>
  imputate_na(p_covered, method = "mice", seed = 123)
```

```{r}
# Plot showing the results of our imputation
mice_na_imp_p_covered |>
  plot()
```
